Typed Functional Genetic Programming

### 2013-07-03 tex/thesis.tex
[tex]

@@

\newcommand{\setDots}[2]{ 
	\lbrace #1 \dots #2 \rbrace
}

@@

@@

	Let $V$ be set of {\it variable names}.  \\* 
	Let $C$ be set of {\it constant names}.	 \\*		
	Then $\Lambda$ is set of {\it \lterms} inductively defined as follows:

	\begin{align*}
		 x   \in V \cup C          &\Rightarrow     x   \in \Lambda \\
		 M,N \in \Lambda           &\Rightarrow ( M N ) \in \Lambda \\
		 x   \in V , M \in \Lambda &\Rightarrow ( \lambda x . M ) \in \Lambda 
	\end{align*}

	\textbf{TODO} 
	\begin{itemize} 
		\item TALK ABOUT "parenthesis" conventions (and packing of lambda abstractions).
		\item BETTER SPECIFICATION $V$ is infinite spočetná (?countable)
	\end{itemize}

\section{Conversion to SKI combinators}
\textbf{TODO}

@@

\subsection{"Barendregt-like" inference and grammar rules}
\label{barlike}

@@

Now we will introduce \textit{Inhabitation tree}, structure slightly different from
\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Henk Barendregt.
We can think about Inhabitation tree as about unfolded Inhabitation machine.
The motivation for using Inhabitation trees is belief that it will help us 
reason about generation of \lterms 
of a given type $\sigma$ and with a given context $\Gamma$.

@@

@@

>>>>

\usepackage[ampersand]{easylist}

@@

\newcommand{\setDots}[2]{ 
	\lbrace #1 , \dots , #2 \rbrace
}

@@

Témata co zmínit v úvodu :\\

\begin{easylist}[itemize]
& Začít asi tim že evoluce je cool a že
  tohle bude o tom jak do GP dát typy.
& Říct jak jde text:
  && Začnem popsanim klasickýho GP podle Kozy
  && Vyslovíme definice okolo lambda termu a typu
  && Vyslovíme definici inhabitačního stromu
     kterej je stromovou verzí IM od Barendrechta.
  && Matematicky popíšem jak na to dem my.
  && Popíšem rozšíření zakladního systému na 
     polymorfní.
  && Popíšem to z programátorskýho hlediska
  && Popíšem příklady na kterejch to demonstrujem
  && Srovnáme to s nějakejma přístupama co už 
     existujou 
& Možná se zmínit o tý představě dvou pólu
  induktivním a deduktivním -
  klasickym GP a Dokazovačem, čim složitější
  typovej sytem tim víc se jedná o deduktivní   
  systém a blíží se to dokazovači.
& Nějak naťuknout už tady v úvodu jak se to má při
  srovnání s tim jak na to dou jiný (Yu atd)
& Výkřiky co zmínit:
  && Na co zaměřený:
     &&& Proof of a concept
     &&& Pořádně to vymáchat v typový teorii
     &&& Spíš než bejt další z řady nástroju
         se snaží zobecnovat ty už existující
     &&& Spíš jednoduchost
         &&&& Ve většině aspektú to sleduje
              Kozu - tzn i v málo řídících
              parametru
     &&& Duraz na generování termu (!)              
     &&& Po implementační stránce
         &&&& Sranda s JS překladem
         &&&& Servrová architektura
         &&&& oboje spíš srandy
  && Na co nezaměřený (ale snad taky ok v tom):
     &&& Performance   
     &&& Mature knihovna co by šla použít robustně
         v cizim kódu - spíš to počítá že to 
         bude dál rozvíjeno a tohle je mezizastávka
\end{easylist}

@@

\newcommand{\then}{\Rightarrow\xspace}

Let $V$ be infinite countable set of {\it 
variable names}.  \\* 
Let $C$ be set of {\it constant names}.	 \\*		
Then $\Lambda$ is set of {\it \lterms} inductively defined as follows:

\newcommand{\lam}[2]{( \lambda #1 . #2 )}

\begin{align*}
x   \in V \cup C          &\then x          \in \Lambda \\
M,N \in \Lambda           &\then (M~N)      \in \Lambda \\
x   \in V , M \in \Lambda &\then \lam{x}{M} \in \Lambda 
\end{align*}
\\
It is usual to use $V = \Sigma^+$ set of all non-empty finite strings of symbols 
from $\Sigma$ where $\Sigma$ is some alphabet set, e.g.  
$
\Sigma =
\setDots{a}{z} \cup 
\setDots{_0}{_9} $.

And $C$ is set supplied by "the user" to enrich 
the language with constant names standing
for some predefined behavior.

A constant may stand just for another \lterm
or it may stand for some predefined constant 
such as 0,1,2,... and primitive operations on
them, e.g. addition. 
But the specific implementations 
of these constants will not be ours big concern 
while reasoning about methods for generating 
\lterms .\\

We use upper case symbols such as 
$M,N,M_1,M_2,...,N_1,N_2,...$
to denote arbitrary lambda terms in contrast to
those in lower case such as
$a,b,c,...,x,y,z,$
$f_1,f_2,...,g_1,g_2,...$
denoting variable names, elements of $V$.\\

Examples of \lterms:\\

\begin{easylist}[itemize]
& $x$
& $\lam{x}{x}$
& $(\lam{x}{(x~x)}~\lam{x}{(x~x)})$
& $\lam{x}{\lam{y}{x}}$
& $\lam{f_1}{\lam{f_2}{\lam{x}{((f_1~x)~(f_2~x))}}}$
\end{easylist}~

We use following notation
abbreviations for better readability:

\begin{easylist}[itemize]
& $M_1~M_2~M_3~\dots~M_n$ for 
  $(\dots((M_1~M_2)~M_3)~\dots~M_n)$ 
& $\lam{x_1 x_2 \dots x_n }{M}$ for
  $\lam{x_1}{\lam{x_2}{\dots\lam{x_n}{M}\dots}}$
\end{easylist}~

\subsection{$\beta$-reduction}

\subsection{$\eta$-reduction}

\subsection{Reductions for $c \in C$}
TODO možná vynechat...

\subsection{Normal form}

\subsection{Long normal form}

\textbf{TODO} 
\begin{itemize} 
	\item TALK ABOUT "parenthesis" conventions 
	      (and packing of lambda abstractions).
	\item Popsat ve stručnosti beta redukci,
	      etha redukci, atd.
	      Ale hlavně promluvit krátce o intuici
	      a nějak to napojit na to jak to používám já
	      pak v tom GP.
	\item promluvit zde o long normal form což propojuje 
		  tuhle sekci se sekcí o barenregt-like 	
		  pravidlech \ref{barlike}.
\end{itemize}

@@

%"Barendregt-like"
\subsection{Inference and grammar rules producing 
terms in Long normal form}
\label{barlike}

As is said in \cite{barendregt10} there are two
inference/grammar rules which together generate
simply typed lambda terms in their long normal forms
(from section \ref{deflam}).

@@

Now we will introduce \textit{Inhabitation tree}, structure slightly different from
\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Barendregt.
We can think about Inhabitation tree as about unfolded Inhabitation machine.
The motivation for using Inhabitation trees is belief that it will help us 
reason about generation of \lterms 
of a given type $\sigma$ and with a given context $\Gamma$.

@@

\section{Our approach to crossover}

\subsection{Conversion to SKI combinators}
\textbf{TODO}

@@

Předběžný náhodně uspořádaný body osnovy ..\\

\begin{easylist}[itemize]
& že je to v haskellu
&& Jak moc hrotit vysvětlení haskellu?
& že je to servrově řešený
& architektura evolučního algoritmu 
&& ty tři typový třídy a ta čtvrtá typová třída
& evaluace
&& vychytavka s vyhodnocenim celýho seznamu aby s eto nemuselo furt inicializovat - inicializace vyhodnocovače jednou za generaci 
& hlavní soubory
&& Eva.hs     - Eva monada
&& GP-Core.hs - Evoluční algoritmus
&& GP-Data.hs - Instance tříd z GP-Core
&& IMx.hs - Term generating
&& TTerm.hs   - typed lambda term
&& TTree.hs   - CTT
&& Server.hs  - main serveru
&& Job.hs     - Registrace problemu
&& Heval.hs   - Evaluator 
&& složka Problems
&&& pro každej problém složka 
&&& ta typicky strukturu
&&&& Problem.hs
&&&& Funs.hs
&& Slozka server
&&& index.html
&&& slozka js/Problems
&&& .js pro kazdej problem
\end{easylist}
